---
title: "Mini Data Analysis"
author: "Natalie"
date: "08/10/2021"
output:
  pdf_document: default
  html_document: default
---

1. -¨[x] Install the [`datateachr`](https://github.com/UBC-MDS/datateachr) package by typing the following into your **R terminal**


2. -[x]Load the packages below

```{r}
library(datateachr)
library(tidyverse)
```

3. -[x] Make a repository in the https://github.com/stat545ubc-2021 Organization. You will be working with this repository for the entire data analysis project. You can either make it public, or make it private and add the TA's and Vincenzo as collaborators. 

4. -[x] When you go to submit, submit a URL to your repository to canvas. 


# Task 1.1: Choose Your Favourite Dataset:heart_decoration:

1: My personal dataset, CTmax_swim_Fry

2: steam_games

3: vancouver_trees

4: parking_meters

## 1.2: Explore datasets:mag_right:

```{r}
### EXPLORE HERE ###
CTmax_swim_Fry
```

```{r}
### EXPLORE HERE ###
steam_games
```

```{r}
### EXPLORE HERE ###
vancouver_trees
```

```{r}
### EXPLORE HERE ###
parking_meters
```
## 1.3: Narrow it down to 2


| Dataset name | Description |
| :---        |    :----:   |  
| CTmax_swim_fry| This subset of my overall data collection is interesting because it was a part of the research project that had the most unknowns about it, as this type of experiment is still quite novel. I'm interested in being able to explore more into this dataset to be able to explore results  | 
|steam_games   |   I'm interested in this dataset because I spend some time gaming, but I don't do it enough to warrant buying loads of games, so being able to have many different games with reviews, ratings, and descriptions readily available for analysis is apealing to me.    | 


## The final decison! 

|Dataset name    |Research Question |
| :---        |    :----:   |
| CTmax_swim_fry| What is the relationship between acclimation temperature and dropout temperature     | 
|steam_games   |Is there a relationship between mature content rating and overall game rating?        | 

**Final Choice :** CTmax_swim_fry
 
 
# Task 2: Exploring your dataset

## Introduction

This is an exploration of the dataset CTmax_swim_fry. This dataset contains critical thermal maximum data for stream-type Chinook salmon (*Oncorhynchus tshawytscha*) fry during swim flume trial experiments. 

## 2.1: complete 4 of 8 excercies / 2.2: provide a brief explanation of reasoning 

*1. Plot the distribution of a numeric variable*

*4. Explore the relationship between 2 variables*

*6. Use a boxplot to look at the frequency of different observations within a single variable*

*8. Use a density plot to explore any of your variables*
