---
title: "mini data analysis deliverable 2"
author: "Natalie"
date: "16/10/2021"
output: github_document
 
---
**Housekeeping: load the following packages and dataset**
```{r message=FALSE, warning=FALSE}
library(ggplot2)
library(ggridges)
library(scales)
library(tsibble)
library(tidyverse)
library(rlang)
library(dplyr)

library(readr)
CTmax_swim_Fry <- read_csv("../CTmax_swim_Fry.csv", 
    col_types = cols(`Fish ID` = col_character(),
        `dropout time` = col_time(format = "%H:%M:%S"), 
        `Acclimation temp C` = col_character(), 
        `fork length (mm)` = col_double(),
        `Mass (g)`= col_double(),
        `Dropout temp` = col_double()))
View(CTmax_swim_Fry)
```


# Task 1: Process and summarize data

## 1.1: Reseach Questions

1. Is body mass influenced by acclimation temperature?

2. Is there a relationship between fork length and body mass? I.e. can I predict that certain length fish will have a certain mass. 

3. Can I find out if acclimation temperature has a more significant influence on either dropout time or dropout temperature?

4. Is there a relationship between body mass and dropout temperature?


## 1.2: Summarizing and Graphing

### Question 1: Is body mass influenced by acclimation temperature?

*Summary* 

1. Computing range, mean, median, and mode for body mass across acclimation temperatures will help better summarize the body mass data, as opposed to looking at the data unmanipulated.This summary helps me answer my question as I'm able to see the differences between body mass and their deviation/spread of values within each acclimation treatment.


```{r warning=FALSE}

#Here the mean is computed to look at what the most average fry mass could be expected to be.
CTmax_swim_Fry %>%
  group_by( `Acclimation temp C`) %>%
  summarise(body_mass_mean = mean(`Mass (g)`, na.rm = TRUE))

#Here the range shows the highest and lowest values for each acclimation treatment.
CTmax_swim_Fry %>%
  group_by( `Acclimation temp C`) %>%
  summarise(body_mass_range = range(`Mass (g)`, na.rm = TRUE))

#The median is a little more descriptive than the mean, as the mean will be affected by outliers in the calculation. However, the median is more robust to the outliers in the data that can be present.
CTmax_swim_Fry %>%
  group_by( `Acclimation temp C`) %>%
  summarise(body_mass_median = median(`Mass (g)`, na.rm = TRUE))

#The standard deviation is measured for each acclimation treatment to see how much variance there is in mass measurements.
CTmax_swim_Fry %>%
  group_by( `Acclimation temp C`) %>%
  summarise(body_mass_sd = sd(`Mass (g)`, na.rm = TRUE))



```

*Graph* 

5. Creating a graph with more than one geom layer can help discern both bigger picture visuals along with some detail all in the same graphic. While the boxplot allows someone to focus on the general distribution of the mass data with each given acclimation treatment, the use of geom layer jitter allows the viewer to see the exact distribution of the mass data without risk of overlapping data points. 

This graphic shows general similarity in mass between the 15C, 18C, and 20C acclimation temperatures, and a distinctly smaller mass with the 24C acclimation temperature, hinting that this higher acclimation temperature may be adversely affecting growth in this age class.

```{r warning=FALSE}
ggplot(CTmax_swim_Fry, aes(`Acclimation temp C`, `Mass (g)`))+
  geom_boxplot()+
  geom_jitter(aes(),
            size = 1.5,
            alpha = 0.3)

```

### Question 2: Is there a relationship between fork length and body mass?

*Summary*

3. Creating a categorical variable out of the fork length can help compare the categorical variable made with body mass (further down). If both have variables with the same constraints (based off the median and range), a count can be performed comparing how many fish scored "Low", "Average", or "High" for both fork length and mass.

```{r warning=FALSE}
(Fry_dropout_length <- CTmax_swim_Fry %>% 
    select(-`dropout time`,-`Mass (g)`, -`Acclimation temp C`,-`Dropout temp`)%>%
    mutate(dropout_length_category = case_when(`fork length (mm)` < 32 ~ "Low",
                                 `fork length (mm)` < 44 ~ "Average",
                                 `fork length (mm)` < 55 ~ "High")))



```

*Graph*

7. The goal of plotting mass and fork length is to visualize the measurements of the fish. Here we can see that fish with a larger mass do tend to be longer, though there are some outliers, which could indicate fish that are either better or worse at competing for food than their cohort. Adjusting the alpha is especially important in this case because we are reliant on the points for data interpretation; as opposed to the boxplots with the overlaid points.There is considerable overlap in the middle of the graph, where average measurements are. Using point instead of jitter is relevant to the data as x-values are numeric and represent the fork length, and it's important to show that some fish had the same fork length but different mass.
```{r warning=FALSE}
ggplot(CTmax_swim_Fry, aes(`fork length (mm)`,`Mass (g)` ))+
  geom_point(colour= "red", alpha=0.2, size=2)

```



### Question 3: Can I find out if acclimation temperature has a more significant influence on either dropout time or dropout temperature?

*Summary*

1. I decided to compute range, mean, median, and standard deviation of dropout temperature across acclimation treatments. This would be a first step into looking at variance in the group. After this I would do the same for dropout time; the goal here would be to do some basic exploratory summaries and then perfrom an ANOVA, to see if there is a difference in significance of variance between acclimation groups. For example, if there is significance in the variance between acclimation treatments in terms of temperature but not time, we can infer that acclimation temperature influences temperature more than time.
```{r warning=FALSE}
#Here the mean is computed to look at what temperature the most fry dropped out at
CTmax_swim_Fry %>%
  group_by( `Acclimation temp C`) %>%
  summarise(dropout_temp_mean = mean(`Dropout temp`, na.rm = TRUE))

#Here the range shows the highest and lowest dropout temperatures for each acclimation treatment.
CTmax_swim_Fry %>%
  group_by( `Acclimation temp C`) %>%
  summarise(dropout_temp_range = range(`Dropout temp`, na.rm = TRUE))

#The median is a little more descriptive than the mean, as the mean will be affected by outliers in the calculation. However, the median is more robust to the outliers in the data that can be present.
CTmax_swim_Fry %>%
  group_by( `Acclimation temp C`) %>%
  summarise(dropout_temp_median = median(`Dropout temp`, na.rm = TRUE))

#The standard deviation is measured for each acclimation treatment to see how much variance there is in droput temperatures .
CTmax_swim_Fry %>%
  group_by( `Acclimation temp C`) %>%
  summarise(dropout_temp_sd = sd(`Dropout temp`, na.rm = TRUE))


```

*Graph*

7. Here I've decided to use ridge plots for both dropout temperature and dropout time. Because there is significant overlap between some of the ridges, adjusting alpha transparency is important to be able to see all of the data. I decided to do two ridge plots instead of boxplots as one of my variables is time, which is better represented through a ridge plot. Here we can see that the 24C acclimated fish were dropping out at a quicker time, but at a comparable temperature range, while the other three acclimation groups had a somewhat similar performance when comparing swim time and final temperature. This may be because as per CTmax protocols, the temperature of the flume or bath is equal to that of the acclimation temperature, so the 24C acclimated fish started swimming in 24C water. On the other hand, the 15C, 18C, and 20C acclimated fish started at lower temperatures and had longer swim times to reach the dropout temperatures recorded.
```{r warning=FALSE}
#This is a ridge plot looking at dropout time in relation to acclimation treatment
(ggplot(CTmax_swim_Fry, aes(`dropout time`, `Acclimation temp C`))+
  ggridges::geom_density_ridges(aes(fill=`Acclimation temp C`), alpha=0.5))

#This is a ridge plot looking at dropout temperature in relation to acclimation treatment
(ggplot(CTmax_swim_Fry, aes(`Dropout temp`, `Acclimation temp C`))+
  ggridges::geom_density_ridges(aes(fill=`Acclimation temp C`), alpha=0.5))


```

### Question 4: Is there a relationship between body mass and dropout temperature?

*Summary*

3. Create a categorical variable sorting dropout mass into low, medium, and high. The new categorical variables are informed by the previous summaries of the data done in question 1. All unrelated categories (length, acclimation temperature, etc.) are removed for ease of reading. With this, I can either do a manual comparison of all the dropout times or I am now set up to make a box plot to visualize the dropout temperatures as a function of body mass.
```{r warning=FALSE}
(Fry_dropout_mass <- CTmax_swim_Fry %>% 
    select(-`dropout time`,-`fork length (mm)`, -`Acclimation temp C`)%>%
    mutate(dropout_mass_category = case_when(`Mass (g)` < 0.45 ~ "Low",
                                 `Mass (g)` < 0.7146667 ~ "Average",
                                 `Mass (g)` < 1.5 ~ "High")))
                                
```
*Graph*

5. Create a graph out of summarized variables that has at least two geom layers. By using a box plot with jitter overlaid, we can see a simplified visual of the data as well as the individual data points. This boxplot shows us that there is considerable overlap with dropout temperatures regardless of mass, leaving us to infer that mass may not have a role in predicting dropout temperature. Alpha transparency here is also adjusted due to crowding of the datapoints.
```{r warning=FALSE}
ggplot(Fry_dropout_mass) + geom_boxplot(aes(x = dropout_mass_category, y = `Dropout temp`)) +
  geom_jitter(aes(x = dropout_mass_category, y = `Dropout temp`), alpha=0.5)+
  labs(y = "Dropout temperature (C)", x = "Mass classes (g)") +
  ggtitle("Dropout temperature as a function of mass")

```

## 1.3: Reflection of research questions


### Question 1: Is body mass influenced by acclimation temperature?

I feel like I've made some headway in terms of addressing if body mass is influenced by acclimation temperature. So far, I'm interested in this information because it can give me an idea of how the environment these fish are in can affect their metabolism and resource partitioning. The 15C, 18C, and 20C fish have very similar mass classes, but the 24C fish have a noticeably smaller mass. This tells me that most of the nutrients are being used up for coping with the warm water, while the three other groups of fish are able to divert extra nutrition towards growth.

I'm not sure if there's more refining to do, but I would like to do some sort of extra analysis or model-fitting to see if there is significance in the preliminary work.


### Question 2: Is there a relationship between fork length and body mass? 

I'm having more difficulty with this one, despite it's simplicity. While I'm able to find clearer answers through graphing, I ran into trouble trying to do summaries with it. I think that the creation of categorical data with both numerical values and comparing how many fish have which combination of categories can help with some very basic inference. However, plotting the measurements made it much simpler to see that while there is variability between mass and fork length, fish with a longer fork length tended to have a higher mass. In terms of refinement, I'd like to look at predictability of body mass and fork length given one of these variables. Is the relationship between fork length and body mass linear among fry?


### Question 3: Can I find out if acclimation temperature has a more significant influence on either dropout time or dropout temperature?

This is the question I'm having the most trouble with, mainly because I'm dealing with three different variables, and one of them is of a different class (time instead of numerical). Especially in terms of summarizing the data in a way to help further my understanding of the question. Additionally, I think it would be better if I focused on dropout temperature instead of time. Due to the nature of the experiment, where temperature is increased over a fixed time interval until the last fish drops out, it can quickly be inferred that the fish who reach the highest temperatures also achieve the longest swim times. With this, I'd like to adjust the research question to focus on the relationship between acclimation treatment and dropout temperature.*Which acclimation temperature yielded fish with the highest thermal tolerance?*


### Question 4: Is there a relationship between body mass and dropout temperature?

The initial findings surprised me because I hypothesized that smaller fish would drop out at lower temperatures. This is because they have less muscle and stored energy, so I thought that they would not be able to swim as long or be as tolerant to the temperature increases. While I don't think there's extra refinement that can happen to this specific question, the followup would be to see if fork length can play a role in dropout temperature. I'm curious about this since a longer fish requires fewer tail beats to swim a set distance, and since the swim flume was set to a specific speed, the longer fish did not have to put in as much work as the smaller fish.


# Task 2: Tidy data


## 2.1: Identifying tidy data

I do believe that my data in its base form is tidy.Each column represents one variable of the experiemnt (Fish ID, Dropout temperature, Dropout time, Mass, Fork Length, and Acclimation temperature). Each row is an observation of data corresponding to one individual fish. As a fish dropped out of the trial, time and temperature of dropout was recorded, and then it was measured for mass and fork length.Acclimation temperature was known beforehand as trials are run with groups of fish all belonging to the same acclimation treatment. Finally, each cell contains one value; either the fish's ID, measurement, performance in the swimming trial, or acclimation treatment.


## 2.2: Tidying and untidying data

I'm going to firstly untidy the data. In order to do that, I'm going to pivot wider the Acclimation Temperature column and sort fish mass under that. This means that each new column representing acclimation temperature will have the value of mass, and mass will no longer be contained within its own column, rendering it untidy.This also becomes untidy as there are large gaps in the table with no values. Because fish only belonged to one acclimation group, the subsequent acclimation columns that don't include the individual fish are left blank.
```{r}
(untidy_mass <- CTmax_swim_Fry %>%
    pivot_wider(id_cols = c(`Fish ID`,`dropout time`,`fork length (mm)`, `Dropout temp`), 
                names_from = `Acclimation temp C`,
                values_from = `Mass (g)`))

```

Now, to tidy the data back up again, I'm going to pivot longer and create two columns; one to encompass the acclimation temperature, and another for the mass.

```{r}
(tidy_mass<- untidy_mass%>%
   pivot_longer(cols=c(-`Fish ID`,-`dropout time`,-`fork length (mm)`, -`Dropout temp`),
                names_to="Acclimation temp C",
                values_to = "Mass (g)"))

```


## 2.3: Research questions to take into Milestone 3 

I've decided to continue with the questions *Is body mass influenced by acclimation temperature?* and *Which acclimation temperature yielded fish with the highest thermal tolerance?*
The dataset that I'll be using is CTmax_swim_Fry_tidy, where I've removed the unnecessary columns of data and empty cells.
```{r}
#1. In order to further tidy my data, I'm first going to remove the columns that don't have any relevant variables.
#2. Depending on which question I'm working on, it may be more suitable to arrange my data by dropout temperature then by mass
#3. 
#4.
(CTmax_swim_Fry_tidy<- CTmax_swim_Fry %>%
   select( `Acclimation temp C`, `Mass (g)`, `Dropout temp` )%>%
   arrange(CTmax_swim_Fry, `Dropout temp`))
 
#Unfortunately ran out of time to perform more functions on my data






```


